Introduction to Market Research

### 2023-08-01 Semester II/Research Method/Paper Finale/draft.Rmd
update dulu deh ya

---
title: Optimization of E-Commerce Product Discount Portfolio Using Modified Dynamic
  Spiral Optimization Algorithm
author: "Mohammad Rizka Fadhli"
date: "| Department of Computational Science, Natural Sciences Faculty, \n| Institut
  Teknologi Bandung\n| Jalan Ganesha No. 10 Bandung, West Java, Indonesia\n| Corresponding
  author's email: 20921004@mahasiswa.itb.ac.id\n"
output:
  pdf_document:
    number_sections: no
    toc: no
    fig_width: 7
    fig_height: 4
    fig_caption: yes
    highlight: breezedark
  word_document:
    toc: no
fontsize: 12pt
affiliation: ITB
bibliography: references.bib
csl: academic-psychiatry.csl
linestretch: 1.25
colorlinks: yes
abstract: The development of the digital economy market in Indonesia is growing every
  year. As a result, marketplaces have to compete with each other to get consumers.
  Blibli.com is one of the pioneer marketplace in Indonesia has strategy to provide
  additional discounts on listed products. Product discount portfolio problem is a
  Binary Linear Programming (BLP) problem with a decision variable in the form of
  a binary number that states whether a product is eligible for a discount or not.
  With many decision variables involved, we developed metaheuristic approaches Spiral
  Dynamic Optimization Algorithm (SDOA) to solve the optimization problem. The SDOA
  solution is proven to be more optimal (generates higher revenur) than the existing
  Blibli.com solution by average $9.8 \%$.
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{20921004@mahasiswa.itb.ac.id}
- \fancyfoot[LE,RO]{page \thepage}
---

@@

BliBli.com is one of the pioneer and top marketplace in Indonesia. To win the competition, they use a product discount strategy. Each listed product on the Blibli.com has an initial price defined by the seller. Blibli.com can intervene in its price indirectly by providing discounts, so consumers pay lower than the initial price. A study at an online retailer in China shows an influence between product discounts and consumer purchase behavior such as purchase incident, purchase quantity, and spending. Especially in specific discount percentage ranges [@cny]. The pricing and discount aim to attract consumers to buy products in the marketplace for a certain period. One of the mathematical models that can be used to determine prices is the Price Elasticity Model, a causality model between price and demand. A study comparing the price elasticity model with linear, polynomial, and exponential bases found that the linear model is the most commonly used, and it is easier to interpret [@linear_reg].

@@

Currently, Blibli.com determines which product to give a discount by trial and error from several possible combinations - (this method is similar to Monte Carlo simulation but on a simpler computational scale). In this research, we develop a computational model based on SDOA to solve the discount product portfolio optimization problem. Then we will compare the solution with the existing product portfolio.

@@

Blibli.com has created the following portfolio:

@@

Each portfolio contains 100 product candidates. Blibli.com has the following problem: choosing the right product from the portfolio to generate the maximal revenue with a budget limit of 5 million rupiahs.

@@

1. The existing solution from blibli.com.
1. The exact solution uses the simplex method created using the R programming language.

@@

In this study, the portfolio used is derived from the calculation of the price elasticity study previously conducted by Blibli.com.

@@

The author would like to thank the Blibli.com data analytic team for entrusting its product portfolio data to be used as a case study in developing an SDOA-based optimization algorithm. The author also thanks Professor Kuntjoro for inspiration for the modification of SDOA. The author stated that there was no conflict of interest during this research.

>>>>

---
title: Optimization of E-Commerce Product Discount Portfolio Using Modified Dynamic
  Spiral Optimization Algorithm
author: "Mohammad Rizka Fadhli"
date: "| Department of Computational Science, Natural Sciences Faculty, \n| Institut
  Teknologi Bandung\n| Jalan Ganesha No. 10 Bandung, West Java, Indonesia\n| Corresponding
  author's email: 20921004@mahasiswa.itb.ac.id\n"
output:
  pdf_document:
    number_sections: no
    toc: no
    fig_width: 7
    fig_height: 4
    fig_caption: yes
    highlight: breezedark
  word_document:
    toc: no
fontsize: 12pt
affiliation: ITB
bibliography: references.bib
csl: academic-psychiatry.csl
linestretch: 1.25
colorlinks: yes
abstract: The development of the digital economy market in Indonesia is growing every
  year. As a result, marketplaces have to compete with each other to get consumers.
  abc.com is one of the pioneer marketplace in Indonesia has strategy to provide
  additional discounts on listed products. Product discount portfolio problem is a
  Binary Linear Programming (BLP) problem with a decision variable in the form of
  a binary number that states whether a product is eligible for a discount or not.
  With many decision variables involved, we developed metaheuristic approaches Spiral
  Dynamic Optimization Algorithm (SDOA) to solve the optimization problem. The SDOA
  solution is proven to be more optimal (generates higher revenur) than the existing
  abc.com solution by average $9.8 \%$.
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{20921004@mahasiswa.itb.ac.id}
- \fancyfoot[LE,RO]{page \thepage}
---

@@

abc.com is one of the pioneer and top marketplace in Indonesia. To win the competition, they use a product discount strategy. Each listed product on the abc.com has an initial price defined by the seller. abc.com can intervene in its price indirectly by providing discounts, so consumers pay lower than the initial price. A study at an online retailer in China shows an influence between product discounts and consumer purchase behavior such as purchase incident, purchase quantity, and spending. Especially in specific discount percentage ranges [@cny]. The pricing and discount aim to attract consumers to buy products in the marketplace for a certain period. One of the mathematical models that can be used to determine prices is the Price Elasticity Model, a causality model between price and demand. A study comparing the price elasticity model with linear, polynomial, and exponential bases found that the linear model is the most commonly used, and it is easier to interpret [@linear_reg].

@@

Currently, abc.com determines which product to give a discount by trial and error from several possible combinations - (this method is similar to Monte Carlo simulation but on a simpler computational scale). In this research, we develop a computational model based on SDOA to solve the discount product portfolio optimization problem. Then we will compare the solution with the existing product portfolio.

@@

abc.com has created the following portfolio:

@@

Each portfolio contains 100 product candidates. abc.com has the following problem: choosing the right product from the portfolio to generate the maximal revenue with a budget limit of 5 million rupiahs.

@@

1. The existing solution from abc.com.
1. The exact solution uses the simplex method created using the R programming language.

@@

In this study, the portfolio used is derived from the calculation of the price elasticity study previously conducted by abc.com.

@@

The author would like to thank the abc.com data analytic team for entrusting its product portfolio data to be used as a case study in developing an SDOA-based optimization algorithm. The author also thanks Professor Kuntjoro for inspiration for the modification of SDOA. The author stated that there was no conflict of interest during this research.

### 2023-08-01 Thesis/Proposal Thesis/Thesis Final/Thesis.Rmd
update dulu deh ya

---
subtitle: "Pengembangan <i>Decision Support System</i> untuk Menyelesaikan Masalah Optimisasi pada Pemilihan dan Penentuan Komposisi Bahan Baku untuk Multi-Produk, Multi-Periode, Multi-Supplier, serta Kontrak Pembelian Minimum Selama Satu Tahun"
title: "Presentasi Tesis"
author: "Mohammad Rizka Fadhli<br>Dr. Saladin Uttunggadewa, M.Si, S.Si<br>Dr. Rieske Hadianti, M.Si, S.Si<br>Prof. Dr. Sri Redjeki, M.Si, S.Si<br>Astri Kurniati, ST., M.App.Sc."
institute: "20921004@mahasiswa.itb.ac.id"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    mathjax: default
    lib_dir: libs
    css: [default,"footer-header.css", fc, fc-fonts]
    seal: true
    nature:
      ratio: '18:10'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
    includes:
      after_body: insert-logo.html
---
layout: true

@@

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
setwd("~/209_ITB/Thesis/Proposal Thesis/Thesis Final")
# reference:
# https://www.garrickadenbuie.com/blog/xaringan-tip-logo-all-slides/
```

@@

# AGENDA HARI INI
## Penjelasan Mengenai Luaran pada _Thesis_

### Pengembangan _Decision Support System_ untuk Menyelesaikan Masalah Optimisasi pada Pemilihan dan Penentuan Komposisi Bahan Baku untuk Multi-Produk, Multi-Periode, Multi-Supplier, serta Kontrak Pembelian Minimum Selama Satu Tahun

@@

# JUDUL THESIS

### Pengembangan _Decision Support System_ untuk Menyelesaikan Masalah Optimisasi pada Pemilihan dan Penentuan Komposisi Bahan Baku untuk Multi-Produk, Multi-Periode, Multi-Supplier, serta Kontrak Pembelian Minimum Selama Satu Tahun

@@

# LUARAN THESIS

@@

@@

Untuk pemenuhan bahan baku tersebut, NFI menggunakan prinsip _multiple sourcing_ dengan perjanjian untuk memasoknya dari enam buah _supplier_. Perjanjian kerja sama antar _supplier_ berupa komitmen pembelian sejumlah bahan baku selama setahun ke depan. Spesifikasi bahan baku dan harganya per-ton berbeda-beda antar _supplier_.

@@

@@

@@

1. Pembelian bahan baku harus berasal dari minimal 2 _suppliers_ untuk menjaga keamanan pasokan.
1. Pembelian bahan baku dari _supplier_ harus memenuhi proporsi portofolio yang ditetapkan di awal tahun oleh __NFI__. Angka ini kelak menjadi patokan berapa tonase minimal yang harus __NFI__ beli kepada masing-masing _supplier_ dalam setahun.

@@

@@

1. Parameter dan variabel yang terlibat merupakan suatu nilai pasti.
1. Variabel yang terlibat meliputi:
    - _Binary_ karena melibatkan pengambilan keputusan bahan baku dari _supplier_ mana yang harus dipesan.
    - _Integer_ karena melibatkan angka kuantitas bahan baku yang harus dipesan. 
    - _Continuous_ karena melibatkan persentase bahan baku yang hendak dialokasikan ke produksi _finished goods_ per pekan.
1. Fungsi _objective_ dan _constraints_ masih berupa _linear_.
    - Meminimumkan ongkos pembelian `harga x tonase`.
    - Kapasitas gudang bahan baku dan rencana produksi produk.
    - Kontrak pembelian selama setahun.
    - Pemenuhan _demand_ bahan baku per pekan.
    - dan seterusnya.

@@

@@

@@

@@

@@

\begin{align}
hc_{j=1} + \sum_{k \in \mathfrak{N}} c_k x_k
\end{align}

@@

>>>>

---
subtitle: "Pengembangan <i>Decision Support System</i> untuk Menyelesaikan Masalah Optimisasi pada Pemilihan dan Penentuan Komposisi Bahan Baku untuk Multi-Produk, Multi-Periode, Multi-Supplier, serta Kontrak Pembelian Minimum Selama Satu Tahun"
title: "T E S I S"
author: "Mohammad Rizka Fadhli<br>Dr. Saladin Uttunggadewa<br>Dr. Rieske Hadianti<br>Prof. Dr. Sri Redjeki<br>Astri Kurniati, ST., M.App.Sc."
institute: "20921004@mahasiswa.itb.ac.id"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  xaringan::moon_reader:
    mathjax: default
    lib_dir: libs
    css: [default,"footer-header.css", fc, fc-fonts]
    seal: true
    nature:
      ratio: '18:10'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
    includes:
      after_body: insert-logo.html
---
layout: true

@@

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
setwd("~/209_ITB/Thesis/Proposal Thesis/Thesis Final")
# reference:
# https://www.garrickadenbuie.com/blog/xaringan-tip-logo-all-slides/
# https://garthtarr.github.io/sydney_xaringan/#48
```

@@

# AGENDA HARI INI

## Penjelasan Mengenai Luaran pada Tesis

@@

# JUDUL TESIS

## Pengembangan _Decision Support System_ untuk Menyelesaikan Masalah Optimisasi pada Pemilihan dan Penentuan Komposisi Bahan Baku untuk Multi-Produk, Multi-Periode, Multi-Supplier, serta Kontrak Pembelian Minimum Selama Satu Tahun

@@

# LUARAN TESIS

@@

# _QUICK RECAP_: Presentasi Proposal Tesis

@@

Untuk pemenuhan bahan baku tersebut, NFI menggunakan prinsip _multiple sourcing_ dengan perjanjian untuk memasoknya dari enam buah _supplier_.

> Catatan: __Satu supplier memasok satu jenis bahan baku__.

Perjanjian kerja sama antar _supplier_ berupa komitmen pembelian sejumlah bahan baku selama setahun ke depan. Spesifikasi bahan baku dan harganya per-unit berbeda-beda antar _supplier_.

@@

Jadi walau beda-beda rasa dan warna, __tapi__ ada satu bahan baku utama.

Jelasin bahwa satu _supplier_ itu memasok hanya satu jenis bahan baku saja.

Lalu jelasin kenapa hanya bisa diproduksi oleh satu jenis bahan baku saja? Kenapa diperbolehkan dua atau lebih?

> Perbedaan mendasar antar bahan baku ini adalah pada warnanya. Untuk produk dengan warna-warna muda, hanya bisa oleh bahan baku warna muda. Kalau gelap, bisa semua.

@@

Jelasin dulu siapa itu PPIC alias _production planning and inventory control_

@@

1. Pembelian bahan baku harus berasal dari minimal 2 _suppliers_ untuk menjaga keamanan pasokan.
1. Pembelian bahan baku dari _supplier_ harus memenuhi proporsi portofolio yang ditetapkan di awal tahun oleh __NFI__. Angka ini kelak menjadi patokan berapa kuantitas minimal yang harus __NFI__ beli kepada masing-masing _supplier_ dalam setahun.

@@

ingat aturan 17 hari

@@

1. Parameter dan variabel yang terlibat merupakan suatu nilai pasti.
1. Variabel yang terlibat meliputi:
    - _Binary_ karena melibatkan pengambilan keputusan bahan baku dari _supplier_ mana yang harus dipesan.
    - _Integer_ karena melibatkan angka kuantitas bahan baku yang harus dipesan. 
    - _Continuous_ karena melibatkan persentase bahan baku yang hendak dialokasikan ke produksi _finished goods_ per pekan.
1. Fungsi _objective_ dan _constraints_ masih berupa _linear_.
    - Meminimumkan ongkos pembelian `harga x kuantitas pembelian`.
    - Kapasitas gudang bahan baku.
    - Kontrak pembelian selama setahun.
    - Pemenuhan _demand_ bahan baku per pekan.
    - dan seterusnya.

@@

Data-data ini diperoleh dari NFI.

sebagai contoh $f_{ik}$ didapatkan dari __resep produksi__.

Sementara $g_{ijk}$ itu diturunkan dari data resep produksi dan demand mingguan bahan baku per item.

@@

Misal suatu produk tidak dipilih, maka $a_{ijk}=0$ walau $f_{ik} = 1$ akibatnya $b_{ijk} = 0$ yang bawahnya terpenuhi

Misal suatu produk dipilih, maka $a_{ijk}=1$ misal $f_{ik} = 1$ akibatnya $b_{ijk} \geq 0$ maka diperlukan $\mu$ agar terpenuhi

@@

Pada minggu I:

\begin{align}
  \sum_{k \in N} (z_{0k} + \hat{x}_{1k} + z_{1k}) - D_1 \leq maxcap
\end{align}

$\forall j \in \left\{ 2,3,4 \right\},$

\begin{align}
  \sum_{k \in N} (z_{(j-1)k} + \hat{x}_{(j-1)k}) - \sum_{i \in P_j} b_{ijk} g_{ik} + z_jk \leq maxcap
\end{align}

@@

Kita pakai rumus luas trapesium

$\frac{1}{2} (alas_1 + alas_2) \times tinggi$

@@

\begin{align}
\frac{1}{2} hc \sum_{j \in M} \sum_{k \in \mathfrak{N}} (z_{(j-1)k} + z_{jk} + \hat{x}_{jk})
\end{align}

@@

__DEMO__

### 2023-08-01 kampus merdeka/ITB Factory Network/market_riset/market_riset.Rmd
update dulu deh ya

---
title: "Introduction to Market Research"
subtitle: |
  | P2MS-ITB 
  | TEACHING FACTORY-NETWORK
author: |
  | Mohammad Rizka Fadhli
  | ikanx101.com
date: "`r format(Sys.time(),'%d %B %Y')`"
fontsize: 10pt
output:
  pdf_document:
    number_section: true
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: breezedark
linestretch: 1.25
---

@@

## _TV Campaign Research_

## _Brand Health_

## _Market Entry Research_

## _Service Audit: Mistery Shopping_

>>>>

---
title: "Introduction to Market Research"
subtitle: |
  | Sebuah Kuliah Perkenalan Sederhana Tentang Market Research
  | Dari Sudut Pandang Praktisi
author: |
  | Mohammad Rizka Fadhli
  | ikanx101.com
date: "`r format(Sys.time(),'%d %B %Y')`"
fontsize: 10pt
output:
  pdf_document:
    number_section: true
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: breezedark
linestretch: 1.25
---

@@
